Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-07-13 manuscript/manuscript.Rmd
adds to SDM ms section; adds esa presentation draft

That population models do not scale easily explains, in part, why species distribution models (SDMs) have become a major tool for land managers and conservation planners. SDMs typically rely on easy-to-collect presence/absence data (but see...) that allow researchers to cover large spatial extents (e.g., ). Likewise, predictors in SDMs tend to be remotely sensed products. Thus, it is relatively straightforward to parameterize and project SDMs at landscape and regional scales, the very scales at which land managers work. However, the limitations of SDMs are well known (cite some review). SOME MORE STUFF... It seems that what land managers really want are population models at the landscape and regional scale, if only the ecological community could provide them.

@@

In general, the QBM climate effects are most correlated with climate effects from the growth model at the individual level (Fig. 6). In no case does the QBM statistical model have strong correlations across all three vital rates (Fig. 6). Thus, for each species the QBM is "missing" climate signals associated with at least one vital rate. This has large impacts on predictions of long term population dynamics, as seen in our equilibrium simulations (Fig. 3a-d).

@@

Conclusions
-----------
This work is not a critique of density-structured population models. In some cases and for certain species, population models based on aggregated data may prove useful and unbiased. However, our work here is the first comparison, to our knowledge, of population models based on individual and aggregated data in a climate change context. Our results confirm theoretical arguments [@Clark2011b] and empirical evidence [@Clark2011; @Clark2012] that individual responses are critical to predicting species' responses to climate change. Thus, forecasts from aggregate level models should be viewed with caution and should never be unaccompianed by uncertainty. Given the importance of demographic data and its current difficulty to collect, we advocate for research on modern methods to collect demographic data more efficiently across environmental gradients in space and time.

@@

```{r figure_6, dependson="plot-options", fig.cap="Correlations between QBM and IPM estimates of climate effects. We ignore sizeXclimate interactions since these are not directly comparable across model types. The QBM does not have multiple vital rates, so its values are repeated across panels within each species. Across top panels, 'growth' = growth regression, 'rec' = recruitment regression, 'surv' = survival regression.", fig.width=6, fig.height=6, cache=FALSE, results='hide'}
##  Figure script to look at correlations between IPM and QBM
##    climate effect estimates.

>>>>

That population models do not scale easily explains, in part, why species distribution models (SDMs) have become a major tool for land managers and conservation planners [see @Guisan2005 for a review]. SDMs typically rely on easy-to-collect presence/absence data [but see @Clark2014 for new methods] that allow researchers to cover large spatial extents [e.g., @Maiorano2013]. Likewise, predictors in SDMs tend to be remotely sensed products. Thus, it is relatively straightforward to parameterize and project SDMs at landscape and regional scales, the very scales at which land managers work. However, the limitations of SDMs are well known [@Elith2009]. It seems that what land managers really want are population models at the landscape and regional scale, if only the ecological community could provide them.

@@

In general, the QBM climate effects are most correlated with climate effects from the growth regression at the individual level (Fig. 6). In no case does the QBM statistical model have strong correlations across all three vital rates (Fig. 6). Thus, for each species the QBM is "missing" climate signals associated with at least one vital rate. This has large impacts on predictions of long term population dynamics, as seen in our equilibrium simulations (Fig. 3a-d). The inability of the QBM to separate the sometimes countervailing effects of climate on survival, growth, and recruitment (Fig. 3e-h) results in inaccurate (Fig. 3a-d) and unprecise (Fig. 4) forecasts. The QBM statistical model struggles to explain variation due to climate variables because they can have positive and negative impacts on different vital rates. When this is the case, as it is for all our species to varying degrees (Fig. 3e-h), statistical models of aggregated population responses will fail.

@@

Conclusions
-----------
This work is not a critique of density-structured population models. In some cases and for certain species, population models based on aggregated data may prove useful and unbiased. However, our work here is the first comparison, to our knowledge, of population models based on individual and aggregated forms of the same data in a climate change context. Our results confirm theoretical arguments [@Clark2011b] and empirical evidence [@Clark2011; @Clark2012] that individual responses are critical to predicting species' responses to climate change. Thus, forecasts from aggregate level models should be viewed with caution and should never be unaccompianed by uncertainty. Given the importance of demographic data and its current difficulty to collect, we advocate for research on modern methods to collect demographic data more efficiently across environmental gradients in space and time.

@@

```{r figure_6, dependson="plot-options", fig.cap="Correlations (r) between QBM and IPM estimates of climate effects. We ignore sizeXclimate interactions since these are not directly comparable across model types. The QBM does not have multiple vital rates, so its values are repeated across panels within each species. Across top panels, 'growth' = growth regression, 'rec' = recruitment regression, 'surv' = survival regression.", fig.width=6, fig.height=6, cache=FALSE, results='hide'}
##  Figure script to look at correlations between IPM and QBM
##    climate effect estimates.

### 2015-07-13 manuscript/manuscript.tex
adds to SDM ms section; adds esa presentation draft

That population models do not scale easily explains, in part, why
species distribution models (SDMs) have become a major tool for land
managers and conservation planners. SDMs typically rely on
easy-to-collect presence/absence data (but see\ldots{}) that allow
researchers to cover large spatial extents (e.g., ). Likewise,
predictors in SDMs tend to be remotely sensed products. Thus, it is
relatively straightforward to parameterize and project SDMs at landscape
and regional scales, the very scales at which land managers work.
However, the limitations of SDMs are well known (cite some review). SOME
MORE STUFF\ldots{} It seems that what land managers really want are
population models at the landscape and regional scale, if only the
ecological community could provide them.

@@

In general, the QBM climate effects are most correlated with climate
effects from the growth model at the individual level (Fig. 6). In no
case does the QBM statistical model have strong correlations across all
three vital rates (Fig. 6). Thus, for each species the QBM is
``missing'' climate signals associated with at least one vital rate.
This has large impacts on predictions of long term population dynamics,
as seen in our equilibrium simulations (Fig. 3a-d).

@@

This work is not a critique of density-structured population models. In
some cases and for certain species, population models based on
aggregated data may prove useful and unbiased. However, our work here is
the first comparison, to our knowledge, of population models based on
individual and aggregated data in a climate change context. Our results
confirm theoretical arguments (Clark et al. 2011b) and empirical
evidence (Clark et al. 2011a, 2012) that individual responses are
critical to predicting species' responses to climate change. Thus,
forecasts from aggregate level models should be viewed with caution and
should never be unaccompianed by uncertainty. Given the importance of
demographic data and its current difficulty to collect, we advocate for
research on modern methods to collect demographic data more efficiently
across environmental gradients in space and time.

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_6.pdf}
\caption{Correlations between QBM and IPM estimates of climate effects.
We ignore sizeXclimate interactions since these are not directly
comparable across model types. The QBM does not have multiple vital
rates, so its values are repeated across panels within each species.
Across top panels, `growth' = growth regression, `rec' = recruitment
regression, `surv' = survival regression.}
\end{figure}

@@

@@

@@

@@

>>>>

That population models do not scale easily explains, in part, why
species distribution models (SDMs) have become a major tool for land
managers and conservation planners (see Guisan and Thuiller 2005 for a
review). SDMs typically rely on easy-to-collect presence/absence data
(but see Clark et al. 2014 for new methods) that allow researchers to
cover large spatial extents (e.g., Maiorano et al. 2013). Likewise,
predictors in SDMs tend to be remotely sensed products. Thus, it is
relatively straightforward to parameterize and project SDMs at landscape
and regional scales, the very scales at which land managers work.
However, the limitations of SDMs are well known (Elith and Leathwick
2009). It seems that what land managers really want are population
models at the landscape and regional scale, if only the ecological
community could provide them.

@@

In general, the QBM climate effects are most correlated with climate
effects from the growth regression at the individual level (Fig. 6). In
no case does the QBM statistical model have strong correlations across
all three vital rates (Fig. 6). Thus, for each species the QBM is
``missing'' climate signals associated with at least one vital rate.
This has large impacts on predictions of long term population dynamics,
as seen in our equilibrium simulations (Fig. 3a-d). The inability of the
QBM to separate the sometimes countervailing effects of climate on
survival, growth, and recruitment (Fig. 3e-h) results in inaccurate
(Fig. 3a-d) and unprecise (Fig. 4) forecasts. The QBM statistical model
struggles to explain variation due to climate variables because they can
have positive and negative impacts on different vital rates. When this
is the case, as it is for all our species to varying degrees (Fig.
3e-h), statistical models of aggregated population responses will fail.

@@

This work is not a critique of density-structured population models. In
some cases and for certain species, population models based on
aggregated data may prove useful and unbiased. However, our work here is
the first comparison, to our knowledge, of population models based on
individual and aggregated forms of the same data in a climate change
context. Our results confirm theoretical arguments (Clark et al. 2011b)
and empirical evidence (Clark et al. 2011a, 2012) that individual
responses are critical to predicting species' responses to climate
change. Thus, forecasts from aggregate level models should be viewed
with caution and should never be unaccompianed by uncertainty. Given the
importance of demographic data and its current difficulty to collect, we
advocate for research on modern methods to collect demographic data more
efficiently across environmental gradients in space and time.

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_6.pdf}
\caption{Correlations (r) between QBM and IPM estimates of climate
effects. We ignore sizeXclimate interactions since these are not
directly comparable across model types. The QBM does not have multiple
vital rates, so its values are repeated across panels within each
species. Across top panels, `growth' = growth regression, `rec' =
recruitment regression, `surv' = survival regression.}
\end{figure}

@@

Clark, J. S., A. E. Gelfand, C. W. Woodall, and K. Zhu. 2014. More than
the sum of the parts: Forest climate response from joint species
distribution models. Ecological Applications 24:990--999.

@@

Elith, J., and J. R. Leathwick. 2009. Species Distribution Models:
Ecological Explanation and Prediction Across Space and Time.

@@

Guisan, A., and W. Thuiller. 2005. Predicting species distribution:
Offering more than simple habitat models.

@@

Maiorano, L., R. Cheddadi, N. E. Zimmermann, L. Pellissier, B.
Petitpierre, J. Pottier, H. Laborde, B. I. Hurdu, P. B. Pearman, A.
Psomas, J. S. Singarayer, O. Broennimann, P. Vittoz, A. Dubuis, M. E.
Edwards, H. A. Binney, and A. Guisan. 2013. Building the niche through
time: using 13,000 years of data to predict the effects of climate
change on three tree species in Europe. Global Ecology and Biogeography
22:302--317.
